---
title: |
  <center> S-052: Intermediate and Advanced Statistical Methods </center>
  <center> for Applied Educational Research </center>
  <center> Spring, 2021 </center>
author: "Sarah Brashear"
date: "Last Modified: 2/20/21"
output: html_document
---

###RQ1: What is the relationship between district socioeconomic status and district achievement in US school districts?
###RQ2: What is the causal effect of own-party identification on viewing time for online ads?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Load in packages
```{r packages, warning = FALSE, message = FALSE}
require(foreign)
require(haven)
require(tidyverse)
require(stargazer)
require(car)
```


##Define the local directory
```{r setwd}
setwd("~/Harvard/RSSS/S052/Spring2021/week5")
```


##RQ1 Data: Nonlinear relationships in SEDA Data
```{r importdata1}
dat_seda <- read.dta("SEDA19.dta")
dat_seda$sesavg2 <- dat_seda$sesavgall*dat_seda$sesavgall
dat_seda$sesavg3 <- dat_seda$sesavgall*dat_seda$sesavg2
```


```{r nonlinear}
#Linear Fit
mod1 <- lm(meanavg ~ sesavgall, data = dat_seda, na.action = "na.exclude")
summary(mod1)

dat_seda$mod1_pred <- predict(mod1)

#Setting the plot range
dat_seda$range <- dat_seda$sesavgall < 3 & dat_seda$sesavgall > -4 & dat_seda$meanavg > 1 & dat_seda$meanavg < 10
  #EK: note this doesn't really make difference for the code below, but will make the variable consistent with the Stata version of the variable
dat_seda$range[is.na(dat_seda$range)] <- FALSE

dat_seda %>%
  filter(range == TRUE) %>%
  ggplot(aes(x = sesavgall)) +
  geom_point(aes(y = meanavg), size = 0.1, color = "darkblue") +
  geom_smooth(aes(y = mod1_pred), color = "black") +
  geom_vline(aes(xintercept = 0), color = "red") +
  scale_y_continuous(breaks = 1:10) +
  ylab("District Average Test Score, 2009-2016") +
  xlab("District Socioeconomic Status") +
  ggtitle("Linear Fit")


#Quadratic Fit
mod2 <- lm(meanavg ~ sesavgall + sesavg2, data = dat_seda, na.action = "na.exclude")
summary(mod2)

dat_seda$mod2_pred <- predict(mod2)

dat_seda %>%
  filter(range == TRUE) %>%
  ggplot(aes(x = sesavgall)) +
  geom_point(aes(y = meanavg), size = 0.1, color = "darkblue") +
  geom_smooth(aes(y = mod2_pred), color = "red") +
  geom_vline(aes(xintercept = 0), color = "red") +
  scale_y_continuous(breaks = 1:10) +
  ylab("District Average Test Score, 2009-2016") +
  xlab("District Socioeconomic Status") +
  ggtitle("Quadratic Fit")


#Cubic Fit
mod3 <- lm(meanavg ~ sesavgall + sesavg2 + sesavg3, data = dat_seda, na.action = "na.exclude")
summary(mod3)

dat_seda$mod3_pred <- predict(mod3)

dat_seda %>%
  filter(range == TRUE) %>%
  ggplot(aes(x = sesavgall)) +
  geom_point(aes(y = meanavg), size = 0.1, color = "darkblue") +
  geom_smooth(aes(y = mod3_pred), color = "blue") +
  geom_vline(aes(xintercept = 0), color = "red") +
  scale_y_continuous(breaks = 1:10) +
  ylab("District Average Test Score, 2009-2016") +
  xlab("District Socioeconomic Status") +
  ggtitle("Cubic Fit")


#Model building table
stargazer(mod1, mod2, mod3, type = "text", column.labels = c("Linear", "Quadratic", "Cubic"), title = "Nonlinear relationships between SES and achievement in US districts", star.char = c("*", "**", "***"), star.cutoffs = c(0.05, 0.01, 0.001), notes = c("*p<0.05; **p<0.01; ***p<0.001"), notes.append = F)


dat_seda %>%
  filter(range == TRUE) %>%
  ggplot(aes(x = sesavgall)) +
  geom_point(aes(y = meanavg), size = 0.1, color = "darkblue") +
  geom_smooth(aes(y = mod1_pred, color = "Linear")) +
  geom_smooth(aes(y = mod2_pred, color = "Quadratic")) +
  geom_smooth(aes(y = mod3_pred, color = "Cubic")) +
  geom_vline(aes(xintercept = 0), color = "red") +
  scale_y_continuous(breaks = 1:10) +
  scale_color_discrete(name = "", breaks = c("Linear", "Quadratic", "Cubic")) +
  ylab("District Average Test Score, 2009-2016") +
  xlab("District Socioeconomic Status") +
  ggtitle("AllPlots")
```


##RQ2 Data: Thanks to Luis Campos and Luke Miratrix for this data story
```{r importdata2}
dat_ads <- read_dta("OnlineAds.dta")
```

#Exploratory Data Analysis
```{r explore}
ggplot(dat_ads, aes(x = seconds)) + geom_histogram() + xlab("Viewing time (seconds)") + ylab("Frequency")

dat_ads %>%
  group_by(republic, partyid) %>%
  summarize(avg_seconds = mean(seconds),
            group_count = n(),
            median_seconds = median(seconds))

dat_ads$fourway <- dat_ads %>%
  mutate(lbl1 = ifelse(republic == 0,
                       "Democrat",
                       "Republican"),
         lbl2 = ifelse(partyid == 0,
                       ", no ID",
                       ", ID"),
         outlbl = paste0(lbl1, lbl2)) %>%
  pull() %>%
  as.factor()

dat_ads$fourway <- factor(dat_ads$fourway, levels = levels(dat_ads$fourway)[c(2,1,4,3)]) #reordering to match Stata output

ggplot(dat_ads, aes(x = seconds, y = stat(density)*1000)) + geom_histogram(bins = 43) + facet_wrap(~fourway) + ylab("Percent") + xlab("Viewing time (seconds)") + ggtitle("seconds")

dat_ads %>%
  mutate(fourway = factor(fourway, levels = levels(dat_ads$fourway)[c(1,3,2,4)])) %>%
  ggplot(aes(y = seconds)) + geom_boxplot() + facet_wrap(~fourway, ncol = 4) + theme(axis.text.x = element_blank()) + ylab("Viewing time (seconds)")
```

#Log Transformation
```{r log}
dat_ads$logsec = log(dat_ads$seconds)

ggplot(dat_ads, aes(x = logsec, y = stat(density)*1000)) + geom_histogram(bins = 43) + facet_wrap(~fourway) + ylab("Percent") + ggtitle("logseconds")

dat_ads %>%
  mutate(fourway = factor(fourway, levels = levels(dat_ads$fourway)[c(1,3,2,4)])) %>%
  ggplot(aes(y = logsec)) + geom_boxplot() + facet_wrap(~fourway, ncol = 4) + theme(axis.text.x = element_blank())
```

#Separate and joint estimation of causal effects
```{r causal}
dat_ads$partyid <- as_factor(dat_ads$partyid)
dat_ads$republic <- as_factor(dat_ads$republic)

mod1 <- lm(logsec ~ partyid, data = filter(dat_ads, republic == "Democrat"))
summary(mod1)

mod2 <- lm(logsec ~ partyid, data = filter(dat_ads, republic == "Republican"))
summary(mod2)

mod3 <- lm(logsec ~ partyid*republic, data = dat_ads)
summary(mod3)

stargazer(mod1, mod2, mod3, type = "text", column.labels = c("Dem", "Rep", "Full"), title = "Effect of own-party identification on viewing time in log-seconds", star.char = c("*", "**", "***"), star.cutoffs = c(0.05, 0.01, 0.001), notes = c("*p<0.05; **p<0.01; ***p<0.001"), notes.append = F)


#Re-testing the effect for Republicans in the full model
coef(mod3) # To understand how R identifies coefficients
linearHypothesis(mod3, "partyidInformed + partyidInformed:republicRepublican", test = "F", verbose = T)
  #Note the combined "estimate" is found under "Estimated linear function..."
  #The associated standard error is indirectly given by "Estimated variance of linear..."; you need to take the squareroot to get to the standard error itself
```