---
title: |
  <center> S-052: Intermediate and Advanced Statistical Methods </center>
  <center> for Applied Educational Research </center>
  <center> Spring, 2021 </center>
author: "Sarah Brashear"
date: "Last Modified: 3/21/21"
output: html_document
---

###RQ:  Are there differential attrition rates between females and non-females in a Computer Science Massive Open Online Course?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Load in packages
```{r packages, warning = FALSE, message = FALSE}
require(haven)
require(tidyverse)
require(stargazer)
```


##Define the local directory
```{r setwd}
setwd("~/Harvard/RSSS/S052/Spring2021/week8")
```

##Import Data
```{r importdata}
dat <- read_dta("AttritionCSMOOC.dta")
```


##Custom lifetab function
```{r}
#The lifetable functions designed for R that I found were either very clunky and/or gave contradictory results to Stata because of some philosophical disagreements on the continuity of time. To save some headache, I just wrote a new function that will perfectly recreate the Stata output.
#Do not use this function in non-S052 contexts. I did not vet this, and do not guarantee its accuracy outside of the examples used in S052.
func_S052lifetab <- function(data,
                             timecol,
                             eventcol){
  data <- data.frame(data)
  if(sum(!(as.numeric(sort(unique(data[,timecol]))) %in%
           as.numeric(1:length(unique(data[,timecol]))))) == 0){
    rowprecursor <- 1:length(unique(data[,timecol]))
  }else{
    print("ERROR; TIME VARIABLE NOT INITIALIZED PROPERLY")
    stop()
  }
  
  temptab <- table(data[,timecol],data[,eventcol])
  out <- data.frame(interval = paste(rowprecursor, 2:(length(rowprecursor)+1)))
  out$Beg.Total <- nrow(data)
  out$Beg.Total <- out$Beg.Total - c(0,cumsum(rowSums(temptab))[-length(rowSums(temptab))])
  out$Deaths <- temptab[,2]
  out$Lost <- temptab[,1]
  out$Hazard <- out$Deaths/out$Beg.Total
  out$Haz.Std.Err <- sqrt((1 - out$Hazard)*(out$Hazard)/(out$Beg.Total - out$Deaths))
  out$Survival <- (out$Beg.Total - out$Deaths)/(out$Beg.Total)
  out$Survival <- cumprod(out$Survival)
  out$Sur.Std.Err <- out$Survival*sqrt(cumsum(out$Deaths/(out$Beg.Total*(out$Beg.Total - out$Deaths))))
  out$Cum.Failure <- 1 - out$Survival
  return(out)
}
```


##Step 1: Review of certification rates (Learning Submission 4)
```{r step1}
#Certification rates by gender
mytab <- table(dat$cert, dat$female)
names(dimnames(mytab)) <- c("Certification", "Female")
dimnames(mytab)[[2]] <- c("Non-female", "Female")
dimnames(mytab)[[1]] <- c("Not certified", "Certified")
mytab
prop.table(mytab, 2)

#Logistic regression to evaluate population disparities.
mod <- glm(cert ~ female, data = dat, family = "binomial")
summary(mod)
  #Coefficients are in logits (log-odds units)
```


##Step 2: Life tables (hazard and survival probabilities)
```{r step2}
ltab <- func_S052lifetab(dat, "relweek", "dropped") #hazard and survival probabilities
ltab

#Descriptive data about survival times
  #total observations
  nrow(dat)
  #subjects
  length(unique(dat$userid))
  #failures in single-failure-per-subject data
  sum(dat$dropped)
  #total analysis time at risk and under observation
  sum(dat$relweek)
    #at risk from t = ...
    min(dat$relweek[dat$dropped == 1]) - 1
    #earliest observed entry t = ...
    min(dat$relweek) - 1
    #last observed exit t = ...
    max(dat$relweek[dat$dropped == 1])

  #generate a hazard probability for each participant for their relative week
  #EK: note that this isn't technically required to make the following plots, the way that it is in Stata
dat$hazardp <- ltab$Hazard[dat$relweek]

#Hazard function with connected lines to visualize the pattern
ggplot(ltab, aes(x = 1:nrow(ltab), y = Hazard)) + geom_point() + geom_line() +
  scale_x_continuous(breaks = 1:nrow(ltab)) +
  xlab("Number of Periods in Trial") +
  ylab("Hazard Probability") +
  ggtitle("hazard")

#generate a survival probability for each participant for their relative week
  #EK: note that this isn't technically required to make the following plots, the way that it is in Stata
dat$survivep <- ltab$Survival[dat$relweek]

#A survival function, plotted appropriately as with Kaplan-Meier "drops" in each discrete bin.
ggplot(ltab, aes(x = 1:nrow(ltab), y = Survival)) + geom_point() + geom_step() +
  scale_y_continuous(breaks = seq(0, 1, .1), limits = c(0,1)) +
  scale_x_continuous(breaks = 1:nrow(ltab)) +
  xlab("Relative Week of Last Activity (< 18)") +
  ylab("Survival Probability") +
  ggtitle("survivalp")


#Estimated median survival time
  #EK: note this is a super lazy way of calculating median survival time that's narrowly passable in this context. Do not try to generalize this code to other survival analysis contexts.
min(which(ltab$Survival < 0.5))
```


##Step 3: Start again, convert from a person-level dataset to a person-period dataset
```{r step3}
dat <- read_dta("AttritionCSMOOC.dta")

#Duplicate rows by duration
dat <- dat %>%
  uncount(relweek, .remove = FALSE) #uncount() by default removes the used variable; we need it for later, so set .remove to FALSE

#Count/label each appearance of a row
dat <- dat %>%
  group_by(userid) %>%
  mutate(period = seq_along(userid))

#Create an "exit" variable that indicates that they experienced the "event"
dat$exit <- 0 #start by saying every row did not experience the event
dat$exit <- (dat$period == dat$relweek)*dat$dropped #identify specifically the rows that did experience the event
```


##Step 4: Building the discrete-time hazard model
```{r step4}
dat$period_cat <- as.factor(dat$period) #Create a categorical version of "period" for ease of use

#Model 1
mod1 <- glm(exit ~ 1, data = dat, family = "binomial")
summary(mod1)

#Model 2
mod2 <- glm(exit ~ period_cat, data = dat, family = "binomial")
summary(mod2)
dat$mod2_prob <- predict(mod2, type = "response")

#Model 3
mod3 <- glm(exit ~ period_cat + female, data = dat, family = "binomial")
summary(mod3)
dat$mod3_prob <- predict(mod3, type = "response")
dat$mod3_logit <- predict(mod3)

stargazer(mod1, mod2, mod3, type = "text",
          add.lines = list(c("neg2ll", round(c(mod1$deviance, mod2$deviance, mod3$deviance), 2)),
                           c("r2pseudo", round(unlist(lapply(list(mod1, mod2, mod3), FUN = function(x) 1 - x$deviance/x$null.deviance)), 4))))
```


##Step 5: Graphing fitted discrete-time hazard models
```{r step5}
#Create a friendlier version of the "female" variable so R doesn't get so easily confused
dat$female_cat <- c("non-female", "female")[dat$female + 1]

#Speed up graphing by plotting one dot per week-by-gender category
  #EK: Note this is typically a Stata hack; R can usually get around this by summarizing the info, but "tagging" unique lines is a decent trick to have up one's sleeve so:
dat$weekftag <- !duplicated(paste(dat$period, dat$female))
table(dat$weekftag)

#Fitted hazard probabilities. Weekly indicator variables allow these to equal sample probabilities.
dat %>%
  filter(weekftag) %>%
  ggplot(aes(y = mod2_prob, x = period)) + geom_point() + geom_line() +
  ylab("Hazard Probability") +
  xlab("Relative Week") +
  scale_x_continuous(breaks = 1:max(dat$period)) +
  ggtitle("m2hazardp")

#Fitted hazard probabilities by gender category from Model 3
dat %>%
  filter(weekftag) %>%
  ggplot(aes(y = mod3_prob, x = period, group = female_cat, color = female_cat)) + geom_point() + geom_line() +
  ylab("Hazard Probability") +
  xlab("Relative Week") +
  scale_x_continuous(breaks = 1:max(dat$period)) +
  ggtitle("m3hazardp")

#Fitted harzard logits by gender category from Model 3.
dat %>%
  filter(weekftag) %>%
  ggplot(aes(y = mod3_logit, x = period, group = female_cat, color = female_cat)) + geom_point() + geom_line() +
  ylab("Hazard Probability") +
  xlab("Relative Week") +
  scale_x_continuous(breaks = 1:max(dat$period)) +
  ggtitle("m3hazardp")
```


##Survival probabilities
```{r survprob}
#Obtain and graph fitted survival probabilities from Model 3.  
#A coding hack. Remember that the sum of logs is the log of a product.
  #EK: Again, probably a cleaner way to accomplish this in R, but a nice demonstration of coding

#The log of the within-bin survival probability (NOT the survival probability)
dat$log1minush <- log(1 - dat$mod3_prob)

#This is a trick to get the desired fitted product (sum of logs)
dat <- dat %>%
  group_by(userid) %>%
  mutate(log_surv = cumsum(log1minush))

dat$mod3_survivep <- exp(dat$log_surv)#Now we obtain the fitted survival probabilities as the products of previous within-bin survival probabilities.

#Fitted survival probabilities
dat %>%
  filter(weekftag) %>%
  ggplot(aes(y = mod3_survivep, x = period, group = female_cat, color = female_cat)) + geom_point() + geom_step() +
  ylab("Fitted Survival Probabilities") +
  xlab("Relative Week") +
  scale_x_continuous(breaks = 1:max(dat$period)) +
  scale_y_continuous(breaks = seq(0, 1, .1), limits = c(0,1)) +
  ggtitle("m3survivep")

#If desired, a hack to appropriately plot Kaplan-Meier "drops" in each discrete bin starting from 100%.
bind_rows(dat,
          data.frame(mod3_survivep = 1,
                     period = 0,
                     weekftag = TRUE,
                     female_cat = c("female", "non-female"))) %>%
  filter(weekftag) %>%
  ggplot(aes(y = mod3_survivep, x = period, group = female_cat, color = female_cat)) + geom_point() + geom_step() +
  ylab("Fitted Survival Probabilities") +
  xlab("Relative Week") +
  scale_x_continuous(breaks = 1:max(dat$period)) +
  scale_y_continuous(breaks = seq(0, 1, .1), limits = c(0,1)) +
  ggtitle("m3survivep")

```


