---
title: |
  <center> S-052: Intermediate and Advanced Statistical Methods </center>
  <center> for Applied Educational Research </center>
  <center> Spring, 2021 </center>
author: "Sarah Brashear"
date: "Last Modified: 2/7/21"
output: html_document
---

###RQ: Does financial aid cause students to enroll in college?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Load in packages
```{r packages, warning = FALSE, message = FALSE}
require(foreign)
require(stargazer)
require(survey)
require(car)
require(tidyverse)
```


##Define the local directory
```{r setwd}
setwd("~/Harvard/RSSS/S052/Spring2021/week3")
```


##Load the data: Data from Dynarski (2003) Does Financial Aid Matter?
\ Article: https://www.jstor.org/stable/3132174
\ As discussed by Murnane & Willett (2010) Chapter 8
\ Available at: https://stats.idre.ucla.edu/stata/examples/methods-matter/chapter8/

```{r importdata}
dat <- read.dta("Dynarski2003.dta")
#Can also use the csv version if you prefer, though there are some variable type differences

#Note that to simplify the the following code, rather than tell R each time that the variable yearsr is not a numeric variable, we're just going to create a non-numeric version of the variable yearsr
dat$yearsr_cat <- as.character(dat$yearsr) #a "character" variable type forces R to think of the values as non-numeric; kind of like football player jersey numbers, you can't add them, they're more like "words"/"characters" than numbers


#Establish survey weights where the primary sampling unit is the household ID
  #Eddie's note: the variable wt88 is not actually a probability weight, the way that the Stata code makes it seem... We need to take the reciprocal to get the actual probability:
# dat$probweights <- 1/dat$wt88
# weightdesign_dat <- svydesign(id = ~hhid,
#                               probs = ~probweights,
#                               data = dat)

  #Or you can just treat it like a regular weight, which it is...
weightdesign_dat <- svydesign(id = ~hhid,
                              weights = ~wt88,
                              data = dat)
```


##Review Variables
```{r summaries}
#Treatment (79-81) and control (82 &83) for eligible group (father deceased) vs. ineligible (father not deceased)
table(dat$fatherdec, dat$yearsr)

#Mean college enrollment by age 23, by two-way groupings
dat %>%
  group_by(fatherdec, yearsr) %>%
  summarize(meancoll = mean(coll, na.rm = TRUE))

#For the one-way groupings, you can modify the "group_by" option
dat %>%
  group_by(fatherdec) %>%
  summarize(meancoll = mean(coll, na.rm = TRUE))
```


#Question 1: Interpreting raw means
```{r q1}
#Ignore for now: Fit a weighted regression model to obtain means in each cell

mod1 <- svyglm(coll ~ yearsr_cat*fatherdec, design = weightdesign_dat, data = dat)
summary(mod1)
dat$mod1_preds <- predict(mod1)

#We used the above regression to obtain a graph of probabilities in each cell:
dat %>%
  group_by(yearsr, fatherdec) %>%
  summarize(meancoll = mean(mod1_preds, na.rm = TRUE)) %>%
  ggplot(aes(x = yearsr, y = meancoll, group = fatherdec, col = fatherdec)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = 81.5, col = "red") +
  ylim(0, 1) +
  xlab("Year in which a senior") +
  ylab("Probability of Attending College by Age 23")
```


##Question 2: Effects by group
```{r q2}
#Showing that the offer variable is a simple recoding of the year variable where 79, 80, and 81 are 1 and 82 and 83 are 0.
table(dat$offer, dat$yearsr)

#Using this variable, we can ask whether the policy affected the focal group and also the reference group
mod2_yes <- svyglm(coll ~ offer, design = weightdesign_dat, subset = fatherdec == "Father deceased")
mod2_no <- svyglm(coll ~ offer, design = weightdesign_dat, subset = fatherdec == "Father not deceased")

summary(mod2_yes)
summary(mod2_no)
```


##Question 3: Interaction model
```{r q3}
#Now, the interaction model for the differences in differences
mod3 <- svyglm(coll ~ offer*fatherdec, design = weightdesign_dat)
dat$mod3_preds <- predict(mod3)
summary(mod3)


#Regression Table
stargazer(mod2_yes, mod2_no, mod3, type = "text", column.labels = c("Focal", "Reference", "Contrast"), title = "Modeling achievement and socioeconomic status in MA and TX")
  #Eddie's note: pretty sure this title was a typo, but wouldn't want you R users to feel left out

#Graph of interaction model
dat %>%
  group_by(offer, fatherdec) %>%
  summarize(meancoll = mean(mod3_preds, na.rm = TRUE)) %>%
  ggplot(aes(x = offer, y = meancoll, group = fatherdec, col = fatherdec)) +
  geom_line() +
  geom_point() +
  ylim(0, 1) +
  scale_x_continuous(breaks = c(0, 1), limits = c(-.25, 1.25), labels = c("No", "Yes")) +
  xlab("Offer of Financial Aid") +
  ylab("Probability of Attending College by Age 23")
```


##Question 4: Linear combination test
```{r q4}
#Test whether the effect for the focal group is 0.

linearHypothesis(mod3, "offer + offer:fatherdecFather deceased = 0", test = "F")
```
