---
title: |
  <center> S-052: Intermediate and Advanced Statistical Methods </center>
  <center> for Applied Educational Research </center>
  <center> Spring, 2021 </center>
author: "Sarah Brashear"
date: "Last Modified: 2/28/21"
output: html_document
---

###RQ: Are there differential attrition rates between females and non-females in a Computer Science Massive Open Online Course?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Load in packages
```{r packages, warning = FALSE, message = FALSE}
require(haven)
require(tidyverse)
require(lmtest)
require(sandwich)
```



##Define the local directory
```{r setwd}
setwd("~/Harvard/RSSS/S052/Spring2021/week6")
```

##Import Data
```{r importdata}
dat <- read_dta("AttritionCSMOOC.dta")
```


#Q1: Cross-tabulations enabling calculation of percents, odds, and odds ratios
```{r Q1}
mytab <- table(dat$cert, dat$female)
names(dimnames(mytab)) <- c("Certification", "Female")
dimnames(mytab)[[2]] <- c("Non-female", "Female")
dimnames(mytab)[[1]] <- c("Not certified", "Certified")
mytab
```


#Q2: Fitting a logistic regression model and plotting fitted values
```{r Q2}
mod_logit <- glm(cert ~ female, data = dat, family = "binomial") # Coefficients are in logits (log-odds units)
summary(mod_logit)
```


#Optional:
Pedagogical visualization of the logistic relationship fitted to two points.
A "dial" fitted to a "switch."
Coding notes: Jitter randomly scatters overlaid categorical points that would otherwise be indistinguishable. Higher values create more scatter.
The 'coef(mod_logit)' commands use stored values from the logistic model above so must be run after fitting the logit model.
```{r optional}
dat$preds <- predict(mod_logit, type = "response")
  #Note the following four lines are an attempt to recreate the "circular" jitter that Stata makes, rather than the "rectangular" jitter that R typically makes. Doesn't matter for most applications, but does help make things visually clearer for the following figures.
tempjitter <- cbind(runif(nrow(dat), 0, 0.1), runif(nrow(dat), 0, 360)*3.1416/180)
tempjitter <- cbind(tempjitter[,1]*cos(tempjitter[,2]), tempjitter[,1]*sin(tempjitter[,2]))
dat$x_jitter <- tempjitter[,1]
dat$y_jitter <- tempjitter[,2]

ggplot(dat) +
  geom_point(aes(x = female, y = preds, shape = "Observed Rate"), size = 3, color = "blue") +
  geom_point(aes(x = female + x_jitter, y = cert + y_jitter*0.15), alpha = 0.2, shape = ".") +
  geom_function(aes(color = "Logistic Fit"), fun = function(x) 1/(1 + exp(-(coef(mod_logit)["(Intercept)"] + coef(mod_logit)["female"]*x))), lwd = 1) +
  theme(legend.title = element_blank(), legend.position = "bottom") +
  xlim(c(-5, 5)) + 
  ylab("Certification Rate") + 
  xlab("Nonfemale = 0, Female = 1") +
  ggtitle("logitrange10")


ggplot(dat) +
  geom_point(aes(x = female, y = preds, shape = "Observed Rate"), size = 3, color = "blue") +
  geom_point(aes(x = female + x_jitter*12, y = cert + y_jitter*0.3), alpha = 0.2, shape = ".") +
  geom_function(aes(color = "Logistic Fit"), fun = function(x) 1/(1 + exp(-(coef(mod_logit)["(Intercept)"] + coef(mod_logit)["female"]*x))), lwd = 1) +
  theme(legend.title = element_blank(), legend.position = "bottom") +
  xlim(c(-50, 10)) + 
  ylab("Certification Rate") + 
  xlab("Nonfemale = 0, Female = 1") +
  ggtitle("logitrange60")


ggplot(dat) +
  geom_point(aes(x = female, y = preds, shape = "Observed Rate"), size = 3, color = "blue") +
  geom_point(aes(x = female + x_jitter*12, y = cert + y_jitter*0.3), alpha = 0.2, shape = ".") +
  geom_function(aes(color = "Logistic Fit"), fun = function(x) 1/(1 + exp(-(coef(mod_logit)["(Intercept)"] + coef(mod_logit)["female"]*x))), lwd = 1) +
  geom_smooth(aes(x = female, y = cert, color = "Linear Probability Model"), method = "lm", se = FALSE, fullrange= TRUE) + 
  theme(legend.title = element_blank(), legend.position = "bottom") +
  xlim(c(-50, 10)) + 
  ylab("Certification Rate") + 
  xlab("Nonfemale = 0, Female = 1") +
  ggtitle("logitvslinear")
```


#Q3: Fitting a "linear probability model" without and with heteroskedasticity-robust standard errors
```{r Q3}
#default linear model output
mod_lin <- lm(cert ~ female, data = dat)
summary(mod_lin)

#robust standard errors output
coeftest(mod_lin, vcov = vcovHC(mod_lin, type = "HC1"))
```

